\bf Visual Inference for Significance and Goodness-of-Fit Testing for a Social Network Model

### 2017-10-30 ASApaper/paper-child.Rnw
respond to hh's comments/questions

Some particularly complicated sets of models are those designed to model network change. A \textit{network} is any set of things, such as people, computers, or neurons, that are connected in some way, through social relations, internet connection, or electrical impulses in the brain. We refer to the ``things" in the network as \textit{nodes}, or \textit{actors} in a social network, and the connections as \textit{edges}, or \textit{ties} in a social network. 
Dependencies inherent to the data make network objects particularly difficult to model. Even more challenging is the situation when we go beyond single instances of a network and consider the dynamics of network change between observed instances. This type of modelling for dynamic networks is often performed on social network data, such as for friendship networks among students or the spread of HIV in  sexually active communities.% of men sleeping with other men. 
These models lack the asymptotics required to perform many well-known goodness-of-fit tests, and the maximum likelihood estimation of parameters is so difficult that it can make significance testing significant as well \citep{goldenberg09}. 
\hh{is the previous sentence a quote? I'm not sure that I follow.}

@@

The paper is outlined as follows: Section~\ref{vi} gives a basic overview of visual inference and the lineup protocol. Section 3 provides an introduction to the our models of interest, stochastic actor-oriented models. Sections 4 and 5 discuss problems with traditional significance tests and goodness-of-fit measures for these models and how visual inference methods can be used as solutions. Section 6 outlines our experimental protocol, and we close with a discussion in Section 7.

\hh{where do the results come in?}

@@

Data visualizations are an important component of data analysis, providing a mechanism for discovering patterns in data. Pioneering research by \citet{gelman:2004}, \citet{Bujaetal} and \citet{majumder:2011} provide methods to quantify the significance of discoveries made from visualizations. \citet{Bujaetal} introduced two protocols, the Rorschach and the lineup protocol, which bridge the gulf between traditional statistical inference and exploratory data analysis. %The Rorschach protocol consists of a set of $m$ (usually, $m=20$) plots (called the {\it null plots}) rendered from data that is consistent with a given null model. The Rorschach protocol helps to understand the extent of randomness in the null model. 
Here, we use the lineup protocol. Under this protocol, a plot of the observed data is placed randomly among a set of $m-1$ null plots (where $m =20$, usually), and human observers are then asked to examine the lineup and to identify the most different plot. If an observer identifies the data plot, this is quantifiable evidence against the null hypothesis. \hh{Similarly, an observer has a chance of 1 in $m$ to pick the data plot from the lineup by simply guessing, ie in a situation where the data plot is virtually indistinguishable from the null plots.}

@@

\section{Stochastic Actor-Oriented Models}

@@

After actor $i$ has been given the opportunity to change, it probabilistically chooses one of its current ties, $x_{ij}$, to change. The probability that actor $i$ changes its current tie to actor $j$ is determined by the \textit{objective function} of the model and a random component, $U$. \hh{XXX can you paraphrase the purpose of U in half a sentence?}
%The random component is assumed to have log-Weibull distribution with location parameter $\mu = 0$ and scale parameter $\sigma = 0$ (see \citet{modelsSnijders}), so the probability density function of $U$ is
%\begin{equation}\label{eq:logweibull}
%g(u) = \exp\{-(u + \exp\{-u\})\}.
%\end{equation}
%The objective function, $f_i$, is the driving force of a SAOM. 
Actor $i$ is aiming to maximize the objective function $f_i$ given the current state of the network, $x$ and the node-level covariates, $\mathbf{Z}$, given as:% which has the form
\begin{equation}\label{eq:objective}
f_i(x, \boldsymbol{\beta}, \mathbf{Z}) = \sum_{k = 1}^K \beta_k s_{ik}(x, \mathbf{Z}),
\end{equation}
where $\boldsymbol{\beta} = (\beta_1, \dots, \beta_K)$ are additional model parameters, each associated with some network statistics, $s_{ik}(x, \mathbf{Z})$, $s_{ik}(x, \mathbf{Z})$, calculated with respect to actor $i$. Network statistics range from the simple outdegree, $s_i(x) = \sum_{i\neq j} x_{ij}$, to the more complicated {\it transitive triplets jumping to different covariate}, $s_i(x, \mathbf{Z}) = \sum_{i \neq j \neq h} x_{ij}x_{ih}x_{hj} \cdot \mathbb{I}(z_i = z_h \neq z_j)$. %, plus many more. At last count, in the software we use to fit these models to network data, 
Version 1.2-3 of \texttt{RSiena} \citep{RSiena}, the software used to fit the models here, provides  over 80 possible effects that can be included in the objective function. We discuss these statistics in more detail in Section~\ref{sec:models}.

@@

This probability dictates which edge change is made by the acting node. The acting node can also choose to \textit{not} change at all, in which case $j \equiv i$.

\hh{XXX how is the probability computed in the case that no change is made? Why is the notation $j \equiv i$ used for no change?}

@@

\section{Discussion}

>>>>

Some particularly complicated sets of models are those designed to model network change. A \textit{network} is any set of things, such as people, computers, or neurons, that are connected in some way, through social relations, internet connection, or electrical impulses in the brain. We refer to the ``things" in the network as \textit{nodes}, or \textit{actors} in a social network, and the connections as \textit{edges}, or \textit{ties} in a social network. 
Dependencies inherent to the data make network objects particularly difficult to model. Even more challenging is the situation when we go beyond single instances of a network and consider the dynamics of network change between observed instances. This type of modelling for dynamic networks is often performed on social network data, such as friendship networks among students or the spread of HIV in \st{drug users sharing needles.} %sexually active communities.% of men sleeping with other men. X ST a bit less jarring hopefully. X 
These models lack the asymptotics required to perform many well-known goodness-of-fit tests, and the maximum likelihood estimation of parameters is so difficult that it can make significance testing difficult as well \citep{goldenberg09}. 
\hh{is the previous sentence a quote? I'm not sure that I follow.} \st{it's a paraphrase... shall I make a quote? Also I definitely botched a word. Hopefully makes more sense now? }

@@

The paper is outlined as follows: Section~\ref{vi} gives a basic overview of visual inference and the lineup protocol. Section~\ref{sec:saom} provides an introduction to the our models of interest, stochastic actor-oriented models. Section~\ref{sec:expersetup} details how we define significance testing and goodness of fit procedures for SAOMs through visual inference, and Section~\ref{sec:res} details the results of a visual inference survey of Amazon Mechanical Turk workers. We close with a discussion in Section~\ref{sec:concl}.

\hh{where do the results come in?}\st{whoops, forgot to updat this with the reorg. }

@@

Data visualizations are an important component of data analysis, providing a mechanism for discovering patterns in data. Pioneering research by \citet{gelman:2004}, \citet{Bujaetal} and \citet{majumder:2011} provide methods to quantify the significance of discoveries made from visualizations. \citet{Bujaetal} introduced two protocols, the Rorschach and the lineup protocol, which bridge the gulf between traditional statistical inference and exploratory data analysis. %The Rorschach protocol consists of a set of $m$ (usually, $m=20$) plots (called the {\it null plots}) rendered from data that is consistent with a given null model. The Rorschach protocol helps to understand the extent of randomness in the null model. 
Here, we use the lineup protocol. Under this protocol, a plot of the observed data is placed randomly among a set of $m-1$ null plots (where $m =20$, usually), and human observers are then asked to examine the lineup and to identify the most different plot. If an observer identifies the data plot, this is quantifiable evidence against the null hypothesis. \hh{Since an observer has a chance of 1 in $m$ to pick the data plot from the lineup by simply guessing, i.e. in a situation where the data plot is virtually indistinguishable from the null plots,} \st{the evidence grows in strength with the number of independent observers identifying the data plot.}

@@

\section{Stochastic Actor-Oriented Models}\label{sec:saom}

@@

After actor $i$ has been given the opportunity to change, it probabilistically chooses one of its current ties, $x_{ij}$, to change. The probability that actor $i$ changes its current tie to actor $j$ is determined by the \textit{objective function} of the model and a random component, $U$, \st{which can be thought of as encompassing any other factors that may be influencing the changes the node makes not accounted for by the parameters in the model.} \hh{XXX can you paraphrase the purpose of U in half a sentence?}
%The random component is assumed to have log-Weibull distribution with location parameter $\mu = 0$ and scale parameter $\sigma = 0$ (see \citet{modelsSnijders}), so the probability density function of $U$ is
%\begin{equation}\label{eq:logweibull}
%g(u) = \exp\{-(u + \exp\{-u\})\}.
%\end{equation}
%The objective function, $f_i$, is the driving force of a SAOM. 
Actor $i$ is aiming to maximize the objective function $f_i$ given the current state of the network, $x$ and the node-level covariates, $\mathbf{Z}$, given as:% which has the form
\begin{equation}\label{eq:objective}
f_i(x, \boldsymbol{\beta}, \mathbf{Z}) = \sum_{k = 1}^K \beta_k s_{ik}(x, \mathbf{Z}),
\end{equation}
where $\boldsymbol{\beta} = (\beta_1, \dots, \beta_K)$ are additional model parameters, each associated with some network statistics, $s_{ik}(x, \mathbf{Z})$, $s_{ik}(x, \mathbf{Z})$, calculated with respect to actor $i$. Network statistics range from the simple outdegree, $s_i(x) = \sum_{i\neq j} x_{ij}$, to the more complicated {\it transitive triplets jumping to different covariate}, $s_i(x, \mathbf{Z}) = \sum_{i \neq j \neq h} x_{ij}x_{ih}x_{hj} \cdot \mathbb{I}(z_i = z_h \neq z_j)$. %, plus many more. At last count, in the software we use to fit these models to network data, 
Version 1.2-3 of \texttt{RSiena} \citep{RSiena}, the software used to fit the models here, provides  over 80 possible effects that can be included in the objective function. We discuss these statistics in more detail in Section~\ref{sec:models}.

@@

This probability dictates which edge change is made by the acting node. The acting node can also choose to \textit{not} change at all. This occurs when the numerator, as calculated for the current state of the network, is larger than for any changes that could be made.

\hh{XXX how is the probability computed in the case that no change is made? Why is the notation $j \equiv i$ used for no change?}\st{that's more for the computation. I'll adjust accordingly}

@@

\section{Discussion}\label{sec:concl}
